---
title: "SEC Edgar Data and Natural Language Processing"
date: "Last updated: `r Sys.Date()`"
output: github_document
---

## Preliminary Work: Install/Load Packages

To try and ensure that this R Notebook will run successfully, we'll use the [renv package](https://cran.r-project.org/web/packages/renv/index.html) to create a project-specific library of packages. This will allow us to install the packages that we need for this project without affecting any other projects that we may be working on. Additionally, the project library will track the specific versions of the dependency packages so that any updates to those packages will not break this project.

The code chunk below will first install the renv package if it is not already installed. Then we will load the package. Next, we'll use the `restore()` function to install any packages listed in the renv.lock file. Once these packages are installed, we can load them into the R session using the `library()` commands. Below the code chunk, we'll list out the packages that will be used in the project demo. And if you run into any trouble using renv, then you can use the second code chunk below and that should be an even more reliable approach to install the required packages.

```{r setup, results='hide', message=FALSE}
# Install renv package if not already installed
if(!"renv" %in% installed.packages()[,"Package"]) install.packages("renv")
# Load renv package
library(renv)
# Use restore() to install any packages listed in the renv.lock file
renv::restore(clean=TRUE, lockfile="../renv.lock")
# Load in the packages
library(edgar)
library(rjson)
```

* The [edgar package](https://cran.r-project.org/package=edgar) wraps the SEC's EDGAR API and provides tools for downloading company filings and conducting sentiment analysis.
* The [rjson package](https://cran.r-project.org/package=rjson) is used to parse the JSON file that maps tickers to CIKs.
* The [rmarkdown package](https://cran.r-project.org/package=rmarkdown) is used to generate this R Notebook.

Since the rmarkdown functionality is built into RStudio, this one is automatically loaded when we open the RStudio. So no need to use the `library()` function for this one. Another observation to make about the code chunk above is that it is labeled as 'setup', which is a special name, which the R Notebook will recognize and automatically run prior to running any other code chunk. This is useful for loading in packages and setting up other global options that will be used throughout the notebook. 

Then if you wish to try and update the versions of the various R packages in the lock file, you can use the `renv::update()` function to update the packages in the project library. However, it is possible that these updates could break the code in this notebook. If so, you may need to adapt the code to work with the updated packages.

My recommendation is to first run through the code using the versions of the packages in the lock file. Then if you want to try and update the packages, you can do so and then run through the code again to see if it still works. If not, you can always revert back to the lock file versions using the `renv::restore()` function.

If you update the packages and get everything working successfully, then you can update the lock file using the `renv::snapshot()` function. This will update the lock file with the versions of the packages that are currently installed in the project library. Then you can commit the updated lock file to the repository so that others can use the updated versions of the packages.

## Setting the User Agent

Before making any requests to the SEC's EDGAR server, we need to configure a user agent string. The SEC's [fair access policy](https://www.sec.gov/os/accessing-edgar-data) requires that all programmatic access to EDGAR identify the requester so that the SEC can contact you if there are any issues with your data requests. Starting with version 2.0.8 of the edgar package, this is enforced as a required parameter for all functions that download data from SEC servers.

The user agent should follow the format `"YourName YourEmail@domain.com"`. Replace the placeholder values in the code chunk below with your own name and a valid contact email address before running any of the code chunks that communicate with the SEC's servers.

```{r useragent}
# Set user agent for SEC EDGAR API access
# Replace with your own name and contact email address
# See https://www.sec.gov/os/accessing-edgar-data for more information
useragent = "YourName YourEmail@domain.com"
```

This `useragent` variable will be passed as a parameter to the edgar package functions that communicate with the SEC's servers throughout the rest of this notebook.

### Alternative Package Installation Code

If you run into any trouble using renv in the code chunk above, then you can use the code chunk below to install the required packages for this analysis. This method will first check if you have already installed the packages. If any are missing, it will then install them. Then it will load the packages into the R session. A potential flaw in this approach compared to using renv is that it will simply install the latest versions of the packages, which could potentially break some of the code in this notebook if any of the updates aren't backwards compatible. 

As long as you have downloaded the entire project repository, the renv chunk above will likely be managing the packages. Thus, the `eval=FALSE` option is used to prevent this chunk from running unless manually executed. So if you only downloaded this one Rmd file, this code chunk should take care of installing the packages for you.

```{r setup2, results='hide', message=FALSE, eval=FALSE}
# Create list of packages needed for this exercise
list.of.packages = c("edgar","rjson","rmarkdown")
# Check if any have not yet been installed
new.packages = list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
# If any need to be installed, install them
if(length(new.packages)) install.packages(new.packages)
# Load in the packages
library(edgar)
library(rjson)
```


## Explore Master Indexes

The first tool from the edgar package that we'll review is the `getMasterIndex()` function. This will create a sub-folder in the R Notebook directory named 'edgar_MasterIndex'. Within this folder, it will download a separate file for each of the four quarters within the requested years and store them in a compressed format (using [gzip](https://www.gzip.org/), which saves the compressed files with a .gz extension). Then for each year, the tool will also aggregate the four quarters of filing data into an R data file (extension .Rda). Below, we download the master indexes file for 2023.

```{r downloadmaster}
getMasterIndex(2023, useragent)
```

We can load the master index from the .Rda file that was saved during the chunk above. One drawback of the .Rda data type is that we cannot assign it a new variable name when loading. So upon running the `base::load()` function, we can look to the Environment tab to see that the stored variable name is `year.master`. Then let's display the first few lines to see the data structure. *Note: we must include the `base::` before the `load()` function because the renv package also has a function named `load()`, which overrides the base R function.*

```{r displaymaster}
base::load("edgar_MasterIndex/2023master.Rda")
head(year.master)
```

If we examine the format of the variables, we can see that the edgar package just stores each variable as a character array, except for the quarter, which is stored as an integer.

```{r classmaster}
class(year.master$cik)
class(year.master$company.name)
class(year.master$form.type)
class(year.master$date.filed)
class(year.master$edgar.link)
class(year.master$quarter)
```

To more efficiently store this data, we can reformat variables that have a finite number of categories as factor variables. Additionally, we can reformat the date variable to a date format.

```{r cleanmaster}
year.master$cik = as.factor(year.master$cik)
year.master$company.name = as.factor(year.master$company.name)
year.master$form.type = as.factor(year.master$form.type)
year.master$quarter = as.factor(year.master$quarter)
year.master$date.filed = as.Date(year.master$date.filed)
```

Then we can save the reformatted master index table to a new name so we can compare the file sizes.

```{r savemaster}
save(year.master,file="edgar_MasterIndex/2023master_new.Rda")
```

Using the `file.size()` function to display the size of the original master index table to the new formatted version, we can see that it reduced in size by over 10%.

```{r sizecomp}
oldsize = file.size("edgar_MasterIndex/2023master.Rda")
oldsize
newsize = file.size("edgar_MasterIndex/2023master_new.Rda")
newsize
newsize/oldsize
```

Since our new table is better formatted, let's replace the old one. But let's also save the old one just in case these formatting changes break some of the other functions in the package.

```{r replacemaster}
file.rename("edgar_MasterIndex/2023master.Rda","edgar_MasterIndex/2023master_old.Rda")
file.rename("edgar_MasterIndex/2023master_new.Rda","edgar_MasterIndex/2023master.Rda")
```


## Searching by CIK

The SEC uses a [Central Index Key (CIK)](https://www.sec.gov/cgi-bin/browse-edgar?action=getcompany) to identify corporations and individuals that file disclosures. You can search the [EDGAR full-text search](https://efts.sec.gov/LATEST/search-index?q=%22central+index+key%22&dateRange=custom&startdt=2021-01-01) to look up a CIK for a specific company; however, it is also possible to look up CIKs programmatically. The SEC publishes a JSON file that maps ticker symbols to CIK numbers, which we can download and parse using the `fromJSON()` function from the rjson package.

```{r tickermap}
# Download the company ticker-to-CIK mapping from SEC EDGAR
ticker.json = fromJSON(file="https://www.sec.gov/files/company_tickers.json")
# Convert the nested list to a data frame
ticker.df = data.frame(
  cik = sapply(ticker.json, function(x) x$cik_str),
  ticker = sapply(ticker.json, function(x) x$ticker),
  company.name = sapply(ticker.json, function(x) x$title),
  stringsAsFactors = FALSE
)
head(ticker.df)
```

Now we can look up the CIKs for any set of ticker symbols. For our example, we'll examine Airbnb Inc. (ABNB) and Las Vegas Sands Corp. (LVS).

```{r tickerlookup}
# Look up CIKs for our target tickers
target.tickers = c("ABNB", "LVS")
target.rows = ticker.df[ticker.df$ticker %in% target.tickers, ]
target.rows
```

```{r cik}
ciks = target.rows$cik
ciks
```


## Downloading Filings

Now that we have the CIK numbers for our target companies, we can use the `getFilings()` function to download the actual text content of their SEC filings. This function creates a subdirectory called `edgar_Filings` in the working directory and saves each filing as a text file. The files are organized in subdirectories by CIK number and form type. Here we download the 10-K annual reports filed by our two companies in 2023.

```{r downloadfilings}
getFilings(cik.no=ciks, form.type='10-K', datebeg='2023-01-01', dateend='2023-12-31', useragent)
```

We can confirm the downloads by looking at the directory structure that was created.

```{r listfilings}
list.files("edgar_Filings", recursive=TRUE) |> head(20)
```


## Parsing Filing Headers

Each SEC filing includes a structured header section that contains metadata about the filer and the submission. The `getFilingHeader()` function extracts this information and returns it as a data frame with columns for attributes like the company's SIC industry code, state of incorporation, fiscal year end, and contact addresses. This structured metadata is useful for filtering or grouping companies in broader analyses.

```{r filingheader}
headerdf = getFilingHeader(cik.no=ciks, form.type='10-K', datebeg='2023-01-01', dateend='2023-12-31', useragent)
head(headerdf)
```

We can look at the column names to understand what information is available in the filing headers.

```{r filingheadernames}
names(headerdf)
```


## Searching Filings for Keywords

The `searchFilings()` function lets us search through the full text of the downloaded filings for specific keywords and count how often each appears. This is useful for quantifying how prominently a company discusses certain topics (such as risk factors, revenue drivers, or competitive dynamics) in its filings, without performing a full sentiment analysis. Below, we search for three keywords across the downloaded 10-K filings.

```{r searchfilings}
searchresult = searchFilings(cik.no=ciks, form.type='10-K', datebeg='2023-01-01', dateend='2023-12-31', keyword=c("risk","revenue","competition"), useragent)
searchresult
```


## Computing Sentiment Measures

Now that we have the CIK identifiers for the companies, we can use the `getSentiment()` function from the edgar package to download the specified filings and generate a table showing some basic textual analysis statistics for the respective filings.

```{r computesentiment}
sentidf = getSentiment(cik.no=ciks, form.type='10-K', filing.year=2023, useragent)
sentidf
```

